Structural replication of nanoporous media using procedural noise

### 2017-02-05 ms.tex
thing
smaating

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract} 
In this paper, we aim to develop a method for producing statistically correct porous media using noise functions. In short, the proposed model is a binary function that returns either true if the input point is inside pore space, or false if within a wall. By using procedural noise, we are not restricted to generating media within a limited box, as the noise functions is continuous on all scales. It is therefore possible to generate geometries at arbitrary scales. In addition, this method is extremely fast and low memory footprints, requiring very little computational power. After presenting such a model, we choose a method to measure the structural properties of nanoporous media that are created either through regular methods (e.g. molecular dynamics) or with our method. The pair distribution function $g(r)$ is chosen as a measure of the structural properties to compare the simulated geometry to our model. Using this measure, we define a full likelihood framework for estimating model parameters given any data set, and continue by validating our own framework by analysing procedural noise data sets with known input parameters. Finally, we generate a nanoporous media with an expand-and-quench technique using molecular dynamics and show that our model correctly reproduces several of the internal properties of the data set, including $g(r)$, porosity and surface area.  \todo{Abstract boer tenkes noe gjennom: faa frem hovedpoenger.}

@@

Procedural generation of physical structures have several advantages over conventional simulations. For example, when producing simulations of nanoporous silica, we are required to calculate and propagate the physical properties of each particle for every time-step throughout the simulation, using a computational framework such as LAMMPS \cite{plimpton1995fast}. This is both time-consuming, and the computational cost can sometimes be be quite large. In addition, the simulation is constrained to a box, so particles are not allowed to move outside the defined boundaries. On the other hand, using procedural noise to carve out nanoporous media is both extremely fast and computationally cost-efficient, in addition to being continuous on all scales. This means that it is possible to simulate 
\todo{Nevn finite size effects}
%In addition, by creating a model that yields properties that are statistically consistent with observations, we are ar

@@

We do not claim that procedural noise methods completely reproduces all structural properties of nanoporous silica, but show that our approach can be used to efficiently create materials that are structurally similar to simulations. Naturally, this all depends on the statistics used for comparing the two models, which in this case is based on the pair correlation function $g(r)$. 
\todo{Nevne at kjemi naturligvis ikke vil være riktig siden antall naboer per atom ikke er riktig. }
%\section{Method}
\section{Procedural content generation}
By procedural content generation, we generally mean a mathematical algorithm that uses limited input data to create complex, sometimes infinite repeating patterns. A firm example would be fractals, such as the Julaia/Mandelbrot [ref mandelbrot] set, where one tests whether the complex quadratic recurrence equation $z_{n+1} = z_n^2 +C$ for $z \in \mathbb C$ converges for each position in a 2D complex plane. With this seemingly simple equation, the level of complexity generated on the boundary between convergence and divergence is infinite, with self-similar patterns repeating themselves to arbitrary detail.
\begin{figure}
\includegraphics[width=.5\textwidth]{orphancity.png}
\caption{Example of city building using procedural methods: Landscape, city infrastructure, buildings and all textures are all automatically generated using procedural methods. Created by one of the authors. \todo{Denne er jo med bare for moro skyld, boer jo fjernes} }
\label{fig:orphancity}
\end{figure}

@@

Procedural algorithmic methods for creating realistic-looking
fractal-like patterns have been around for more than 30 years. The most famous of these
algorithms is Perlin noise, first developed by Ken Perlin in 1983 for
use in the movie ``Tron''\cite{perlin1985image}.

Perlin noise is a type of pseudo-random gradient noise that mimics a Gaussian field, but at low computational cost. An improved version of Perlin noise is called simplex noise, which has several advantages over standard Perlin noise in terms of scalability and speed \cite{perlin:2002}. Perlin noise was developed in response to the lack of detail 1980's contemporary graphics, where polygonal surfaces often were bland and left without any detail. In response to this problem, Perlin noise was developed in order to continuously perturb the light bouncing off flat surfaces, effectively yielding bump mapping and producing a much more realistic image.

An overview of the current state of procedural noise functions can be
found in \cite{lagae:2010}. Perlin noise is a band limited repeatable pseudo-random function
from $\mathbb R^n \to \mathbb R$ that is computationally fast, and has
properties which make it suitable for creating patterns that mimic
nature. Perlin noise is an approximation to Gaussian filtered noise
implemented as a pseudo-random spline. The algorithm uses a predefined
grid of gradients pointing in a random direction, where for any point
within a grid, we interpolate the gradients from the four corners. The
algorithm is as follows:
\begin{itemize}
  \item[1.] For a point, obtain the four closest gradients in the grid.
   \item [2.] For each of the gradients, calculate the dot product
     between the gradient and a vector defined by the  the distance
     between the the point and the current grid corner. 
   \item [3.] Instead of linear interpolation $F(t) = t$, use a fade function to produce smoother interpolated values. Perlin noise typically uses the following fade function: $F(t) = t ^3 (t  (6 - 15) + 10)$.
    \item [4.] Obtain the final value by linearly interpolating between the x-axis and
      use this result to linearly interpolate the y-axis. 
\end{itemize} 
Simplex noise works in a similar manner, but works on on simplices instead of hypergrids, and so requires fewer computational steps, scales better in higher ($\ge 3$) dimensions and has fewer anisotropic properties.

\subsection{Properties of Perlin/simplex noise}
\begin{figure}
\includegraphics[width=.5\textwidth]{noise.eps}
\caption{Perlin noise (left) versus simplex noise (right). Note the
  straight x-y pattern in the single upper Perlin mode and the less
  anisotropic simplex mode. Lower part: combining octaves tend
  to remove these patterns.}
\label{fig:noise}
\end{figure}

@@

We now perform a parameter fitting analysis on realistic simulated nanoporous media. Using LAMMPS \cite{plimpton1995fast}, we simulate a $100^3\AA$  box of 100K SiO2 silica particles at an initial temperature of $4500 K$, using the Vashishta-potential. After an initial period, we expand the box to $159^3 \AA$ and start cooling until quenching is achieved at about $300K$. The resulting nanoporous media is depicted on the right side in figure \ref{fig:porous_vs_model}, with a porosity of $75\%$. \todo{Skrive mer om lagingen?}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract} 
In this paper, we aim to develop a method for producing statistically correct simulations of nanoporous media without the use of conventional Molecular Dynamics Simulator (MDS). In short, the proposed model is a binary function that returns either true if the input position is in a void (pore), or false if within a wall. By using procedural noise, we are not restricted to generating media within a limited box, as the noise functions is continuous on all scales. It is therefore possible to simulate nanoporous media of arbitrary size. In addition, this method is extremely fast, requiring very little computational power to calculate whether a particle is within a pore or not. After presenting such a model, we select a method to measure the structural properties of nanoporous media that are created either through regular methods (e.g. LAMMPS) or with our method. The measure we selected is the radial distribution function (RDF) $g(r)$, which is regarded as the de-facto standard method for capturing the structural properties of porous media. Using this measure, we define a full likelihood framework for estimating model parameters given any data set, and continue by validating our own framework by analysing procedural noise data sets with known input parameters. Finally, we perform a full likelihood analysis of silica SiO2 data set simulated in LAMMPS, and show that our model correctly reproduces several of the internal properties of the data set, including $g(r)$, porosity and surface area.  \todo{Abstract boer tenkes noe gjennom: faa frem hovedpoenger.}

@@

Procedural generation of physical structures have several advantages over conventional simulations. For example, when producing simulations of nanoporous silica, we are required to calculate and propagate the physical properties of each particle for every time-step throughout the simulation, using a computational framework such as LAMMPS \cite{plimpton1995fast}. This is both time-consuming, and the computational cost can sometimes be be quite large. In addition, the simulation is constrained to a box, so particles are not allowed to move outside the defined boundaries. On the other hand, using procedural noise to carve out nanoporous media is both extremely fast and computationally cost-efficient, in addition to being continuous on all scales. This means that it is possible to simulate

%In addition, by creating a model that yields properties that are statistically consistent with observations, we are ar

@@

We do not claim that procedural noise methods completely reproduces all structural properties of nanoporous silica, but show that our approach can be used to efficiently create materials that are structurally similar to simulations. Naturally, this all depends on the statistics used for comparing the two models, which in this case is based on the radial distribution function (RDF) $g(r)$.

%\section{Method}
\section{Procedural content generation}
By procedural content generation, we generally mean a mathematical algorithm that uses limited input data to create complex, sometimes infinite repeating patterns. A firm example would be fractals, such as the Julaia/Mandelbrot [ref mandelbrot] set, where one tests whether the complex quadratic recurrence equation $z_{n+1} = z_n^2 +C$ for $z \in \mathbb C$ converges for each position in a 2D complex plane. With this seemingly simple equation, the level of complexity generated on the boundary between convergence and divergence is infinite, with self-similar patterns repeating themselves to arbitrary detail.
\begin{figure}
\includegraphics[width=.5\textwidth]{orphancity.png}
\caption{Example of city building using procedural methods: Landscape, city infrastructure, buildings and all textures are all automatically generated using procedural methods. Created by one of the authors. \todo{Denne er jo med bare for moro skyld, boer jo fjernes} }
\label{fig:orphancity}
\end{figure}

@@

Procedural algorithmic methods for creating realistic-looking
fractal-like patterns have been around for more than 30 years. The most famous of these
algorithms is Perlin noise, first developed by Ken Perlin in 1982 for
use in the movie ``Tron''.

Perlin noise is a type of pseudo-random gradient noise that mimics a Gaussian field, but at low computational cost. An improved version of Perlin noise is called simplex noise, which has several advantages over standard Perlin noise in terms of scalability and speed. Perlin noise was developed in response to the lack of detail 1980's contemporary graphics, where polygonal surfaces often were bland and left without any detail. In response to this problem, Perlin noise was developed in order to continuously perturb the light bouncing off flat surfaces, effectively yielding bump mapping and producing a much more realistic image.

An overview of the current state of procedural noise functions can be
found in \cite{lagae:2010}.  Perlin noise \citep{perlin:2002} 
is a band limited repeatable pseudo-random function
from $\mathbb R^n \to \mathbb R$ that is computationally fast, and has
properties which make it suitable for creating patterns that mimic
nature. Perlin noise is an approximation to Gaussian filtered noise
implemented as a pseudo-random spline. The algorithm uses a predefined
grid of gradients pointing in a random direction, where for any point
within a grid, we interpolate the gradients from the four corners. The
algorithm is as follows:
\begin{itemize}
  \item[1.] For a point, obtain the four closest gradients in the grid.
   \item [2.] For each of the gradients, calculate the dot product
     between the gradient and a vector defined by the  the distance
     between the the point and the current grid corner. 
   \item [3.] Instead of linear interpolation $F(t) = t$, use a fade function to produce smoother interpolated values. Perlin noise typically uses the following fade function: $F(t) = t ^3 (t  (6 - 15) + 10)$.
    \item [4.] Obtain the final value by linearly interpolating between the x-axis and
      use this result to linearly interpolate the y-axis. 
\end{itemize} 
Simplex noise works in a similar manner, but requires fewer
computational steps, scales better in higher ($\ge 3$) dimensions and has fewer 
anisotropic properties.

\subsection{Properties of Perlin/Simplex noise}
\begin{figure}
\includegraphics[width=.5\textwidth]{noise.eps}
\caption{Perlin noise (left) versus simplex noise (right). Note the
  straight x-y pattern in the single upper Perlin mode and the less
  anisotropic simplex mode. Lower part: combining octaves tend
  to remove these patterns.}
\label{fig:noise}
\end{figure}

@@

We now perform a parameter fitting analysis on realistic simulated nanoporous media. Using LAMMPS \cite{plimpton1995fast}, we simulate a $100^3\AA$  box of 65K SiO2 silica particles at an initial temperature of $4500 K$, using the Vashishta-potential. After an initial period, we expand the box to $159^3 \AA$ and start cooling until quenching is achieved at about $300K$. The resulting nanoporous media is depicted on the right side in figure \ref{fig:porous_vs_model}, with a porosity of $75\%$. \todo{Skrive mer om lagingen?}
